---
title: "Capesize - Time Series Analysis"
author: "Sachin Manjhi"
date: "September 6, 2018"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(forecast)
```

## Read C8_14 route freight the data from csv file

```{r input}
# C8 historical data
c8_raw_data <- read.csv(file="c8_rates.csv", stringsAsFactors = FALSE)
head(c8_raw_data)
```

## Convert raw data to Timeseries
```{r convert}
c8_ts_data <- ts(c8_raw_data$c8_14, frequency=12, start=c(2011,1))
c8_ts_data
```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Plot values as time series
```{r plot ts}
plot.ts(c8_ts_data, axes=TRUE, ylab="Freight Rate", main="Freight Rate : C8_14 route")

# Plot the mean value
mean_y <- mean(c8_ts_data)
abline(h=mean_y, lt=21, col='blue')

# plot the regression line
linearFit <- tslm(c8_ts_data ~ trend + season)
abline(linearFit, lt=22, col='green')

```


## Decompose the time series into components
```{r decompose}
c8_ts_components <- decompose(c8_ts_data)
plot(c8_ts_components)
```


## Lets look closely at the random variation
```{r randomVariation}
c8_ts_random <- c8_ts_data - c8_ts_components$seasonal - c8_ts_components$trend
plot(c8_ts_random, main="Random variation")

trend <- c8_ts_components$trend
seasonaity <- c8_ts_components$seasonal
random <- c8_ts_data - c8_ts_components$seasonal - c8_ts_components$trend

ds <- data.frame(c8_ts_data, trend, seasonaity, random)

write.csv2(file = "tscomponents.csv", ds)

```

## Lets try to fit a HoltWinter Model
```{r holtWinter}

c8_18_Forecast_1 <- HoltWinters(c8_ts_data, beta = 0, gamma = 0)  #simple exponential smoothing
plot(c8_18_Forecast_1)

c8_18_Forecast_2 <- HoltWinters(c8_ts_data)                       #triple exponential smoothing
plot(c8_18_Forecast_2)
```

## Lets investigate how well an ARIMA model will do
```{r ACF plot}
acf(c8_ts_data, lag.max=10)  
pacf(c8_ts_data, lag.max=10)
```
# ACF  - We see that the autocorrelations for lags 1, 2, 3 are positive and tail off after lag 3.
# PACF - we see that the partial autocorrelation (pac) at lag 1 is positive and exceeds the significance bounds, while  the partial autocorrelation 
#        at lag 3 is negative and also exceeds the significance bounds.

# Hence Following ARMA models are possible:
# ARMA(0,3), ARMA(3,0), ARMA(3,3), ARMA(3,2) etc.


```{r arima}
c8_arima_model <- auto.arima(c8_ts_data)
c8_arima_model
```


```{r predict}
fcast <- forecast(c8_arima_model, h=30)
plot(fcast)

```


